JWST Noise Floor I: Systematic Error Sources with JWST NIRCam Time Series

### 2020-05-10 ms.tex
updated the crosshatch calculation for latest Full Frame Grism field point

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.2
\documentclass{aastex62}

@@

\begin{figure}[!hbtp]
\centering
\includegraphics[width=.49\columnwidth]{allamps.pdf}
\includegraphics[width=.49\columnwidth]{allamps_long_dark.pdf}
\caption{Reference pixel time series for a two frame integration (Left) and a long dark integration of 108 frames (Right). Each of the plots show that the pre-amplifier resets between frames in an integration cause discontinuities in the reference signal. Amplifiers 1 and 4 track the time all through a frame with the side reference pixels, whereas amplifiers 2 and 3 in the middle only contain information on the bottom and top reference pixels at the beginning and ending of a frame.
The long dark exposure example of 108 integrations shows the behavior over long timescales.}\label{fig:ampResetDark}
\end{figure}

@@

\begin{figure}[!hbtp]
\centering
\includegraphics[width=.79\columnwidth]{amplifier_offsets_in_otis_lw_grism.pdf}
\caption{An example slope image from the LW OTIS Stability test.
{\it First Panel:} No reference pixel correction is applied, leaving offsets at the amplifier boundaries.
{\it Second Panel:} Averaging all the reference pixels within an amplifier and subtracting the result reduces the offsets between amplifiers but there are still noticeable boundaries at X= 512, 1024 and 1536.
{\it Third Panel:} Reference pixel correction using the 4 non-illuminated pixels in each column significantly reduces offsets between amplifiers, but introduces read noise to each column.
{\it Fourth Panel:} Background subtraction with the mean value of the pixels below Y=15 and above Y=190 eliminates offsets between amplifiers without introducing any substantial new noise source.
}\label{fig:ampOffsetsOtisGrismSlope}
\end{figure}

\clearpage
\section{1/f Noise}
The readout circuitry for NIRCam and other JWST NIR instruments introduces correlated noise from pixel-to-pixel.
This is visible when examining the time series of pixels as they are read out within a frame in the order they are addressed.\footnote{See a diagram of the readout directions at \url{https://jwst-docs.stsci.edu/near-infrared-camera/nircam-instrumentation/nircam-detector-overview/nircam-detector-readout}}
The power spectrum of this pixel-by-pixel time series reveals high power at low frequencies that drops approximately like a 1/f power law in the power spectrum of the time series.
Figure \ref{fig:pixelTSeriesPSpec} shows the average Lomb Scargle periodogram from amplifier over 108 dark frames.
To create the periodogram, the reference pixels are used to remove amplifier offsets between frames.
Next, the average of all frames is subtracted to remove the bias and all pixels that deviate by more than 80 DN from the average are masked to eliminate bad pixels.

@@

We note that Figure \ref{fig:genericSourceErrorContrib} shows a somewhat counter-intuitive result: read noise dominates for bright source and becomes less important for faint sources.
For typical observations, such as faint galaxies or distant stars, the opposite is true: read noise is more of a concern for faint sources than bright ones.
So why is there a discrepancy?
The difference in these two science cases is that for exoplanet transit observations, the detector well depth is filled to approximately the same level ($\sim 32,000$ DN or 50\% well depth per pixel) regardless if the system's magnitude is 6 or 11.
For the bright source with a few groups up the ramp, the read noise will be a larger fraction of 32,000 DN than for a faint source with many groups up the ramp.
This is different from a faint galaxy in the $K > 20$ range, which will not fill the detector well.
For a fixed exposure time, a $K=25$ and $K=20$ galaxies would be observed with approximately the same number of groups per integration so the read noise in DN from each source in an integration would be approximately the same whereas the number of DN from the source per integration would be 100 times larger for the $K=20$ galaxy.

@@

Figure \ref{fig:cv3ExtractedSpectrum} shows an extracted spectrum from Cryogenic Vacuum 3 testing where a broadband point source is dispersed by the grism.
The slope image is calculated from the raw ramp using the internal NIRCam ramps-to-slopes pipeline \texttt{nchdas}.
The background is subtracted with a 2nd order polynomial fit and the flux is extracted with an optimally weighted summation \citep[e.g.][]{horne1986optimalE}.
The spectrum is normalized by a 3rd order spline fit with 400 knots to remove the broad shape of the spectrum and the throughput curves.
The resulting spectrum shows high frequency ($\sim$0.2 px$^{-1}$ to $\sim$0.5 px$^{-1}$) variations far in excess of the photon and read noise.
These oscillations increase in frequency at longer wavelengths (X pixels).
The variations have peak-to-valley amplitudes of about 8\%.
However, we expect that the fringing structure will hold relatively constant during time due to the temperature stability of the instrument.

\begin{figure}[!hbtp]
\centering
\includegraphics[width=.99\columnwidth]{fringing_grism_cv3.pdf}
\caption{A CV3 grism spectrum over the time series grism test shows evidence for fringing and where the oscillations exceed the photon and read noise (green point).
An illustrative sinusoidal function is plotted for comparison to better see how the frequencies change as a function of wavelength (X pixel).
}\label{fig:CV3GrismSpecFringing}
\end{figure}

\clearpage
\section{Subpixel Crosshatching}

@@

The crosshatch structure of the detectors extends down to the subpixel level, so it will not be fully corrected with a flat field division.
This sub-pixel structure has been imaged with microscopy on a Euclid HgCdTe detector \citep{shapiro2018crosshatch}.
Atomic force microscopy shows topological features that are approximately 1.2~$\mu$m in width \citep{chang2008surfaceMorphologyHgCdTe} compared to the 18~$\mu$m pixel sizes.
The width of the structures can also be estimated from the behavior for crosshatch lines as they cross pixel boundaries \citep{ninan2019crosshatchHPF}.
We estimate that the crosshatch pattern oriented at 90.9$\degree$ crosses 1 horizontal pixel for every 64 vertical pixels.
While crossing the boundary, there are $\approx 38$ rows where the crosshatch pattern spans more than 2 pixels.
If the crosshatch pattern has a tophat response function, then these 38 rows where the pattern spans two pixels imply a tophat full width of 0.6 pixels or a physical width of 10.8~$\mu$m for an 18~$\mu$m pixel pitch.
This is more than twice the estimate from \citet{ninan2019crosshatchHPF} for an HgCdTe used on the Habitable Planet Finder.
We expect that the crosshatch pattern's width varies among detectors and that a tophat function is a poor approximation of the actual subpixel response.
Within NIRCam detectors, there are large variations in the strength and orientation of crosshatch features.

The crosshatch pattern is wavelength dependent, as seen in Figure \ref{fig:crossHatchWavelengthDep}, where the throughput variations are the largest for short wavelengths (resolving crystal structures) and smallest for the long wavelengths.
This is visible both in the throughput cross section of the flat field and the Fourier amplitude of the crosshatch as a function of wavelength.
We find a steeper wavelength dependence to the crosshatch pattern near 0.9$\degree$ from horizontal in the frequency domain or 0.9$\degree$ from vertical in the length domain.
%We also fit the crosshatch pattern.

@@

For the F300M filter and ALONG detector, we find 
for $\theta_1 = 0.906 \pm 0.001 \degree, \theta_2 = 68.605 \pm 0.001 \degree, \theta_3 = 113.075 \pm 0.001, a_1 = 2.774 \pm 0.002 \times 10^{-7}, a_2 = 1.569 \pm 0.002 \times 10^{-7}, a_3 = 1.744 \pm 0.002 \times 10^{-7}, b=0.345 \pm 0.0005 $ px$^{-1}$, $c=2.05 \pm 0.07 \times 10^{-2}$ px$^{-1}$, $d=6.78 \pm 0.01 10^{-3}$ and $e_b=0.26 \pm 0.01$ px$^{-1}$.
In other words, the separations between the vectors are $67.699 \degree, 67.831 \degree$, and $44.470 \degree$.
The uncertainties were simply derived from the diagonals of the covariance matrix and likely are underestimated.

NIRCam's 10 SCA's have very similar crosshatch angles with separations of 44.5 $\degree$ and 67.7 to 67.8 $\degree$.
They tend to have one primary crosshatch axis oriented either along the X pixels or the Y pixels, with two flanking axes at $\sim 67.7 \degree$ to either side.
The long wavelength arrays have a more pronounced crosshatch than the short wavelength arrays.
The NRCB4 array's primary crosshatch direction is most closely aligned with a pixel axis (the Y direction) than any SCA.
Figure \ref{fig:crosshatchModelGallery} shows the best-fit crosshatch 2D power spectra.

@@

\clearpage
\subsection{Position of SubGrism Array}

@@

The three subarrays considered are shown in Figure \ref{fig:subpixScanSimulation}.
We exclude reference pixels (which form a 4 pixel wide boundary around the detector) and an additional row on the bottom and top of the array because these illuminated rows are outliers in the flat field.
The first position is nearly the same as the SUBGRISM64 subarray position in the grism time series mode, except that the SUBGRISM64 includes the first 5 rows which are excluded here from the crosshatch analysis.
The field point used for the SUBGRISM64 mode is expected to be the same one as will be used for the SUBGRISM128 and SUBGRISM256 subarrays.
The position at the top of the ALONG detector has the smallest amplitude of crosshatch pattern in all three directions.
The best-fit amplitudes are used as input to the jitter simulations in Section \ref{sec:CrosshatchSim}.
We find that the maximum flux change for a 0.1 pixel shift is 118 ppm for the nominal grism position, 108 ppm for the top of the array and 153 ppm at the full frame grism position.
\textcolor{red}{\textbf{Update this for latest full frame grism field point}}

@@

>>>>

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.2
\documentclass[]{aastex62}

@@

\begin{figure*}[!hbtp]
\centering
\includegraphics[width=.49\columnwidth]{allamps.pdf}
\includegraphics[width=.49\columnwidth]{allamps_long_dark.pdf}
\caption{Reference pixel time series for a two frame integration (Left) and a long dark integration of 108 frames (Right). Each of the plots show that the pre-amplifier resets between frames in an integration cause discontinuities in the reference signal. Amplifiers 1 and 4 track the time all through a frame with the side reference pixels, whereas amplifiers 2 and 3 in the middle only contain information on the bottom and top reference pixels at the beginning and ending of a frame.
The long dark exposure example of 108 integrations shows the behavior over long timescales.}\label{fig:ampResetDark}
\end{figure*}

@@

\begin{figure*}[!hbtp]
\centering
\includegraphics[width=.79\columnwidth]{amplifier_offsets_in_otis_lw_grism.pdf}
\caption{An example slope image from the LW OTIS Stability test.
{\it First Panel:} No reference pixel correction is applied, leaving offsets at the amplifier boundaries.
{\it Second Panel:} Averaging all the reference pixels within an amplifier and subtracting the result reduces the offsets between amplifiers but there are still noticeable boundaries at X= 512, 1024 and 1536.
{\it Third Panel:} Reference pixel correction using the 4 non-illuminated pixels in each column significantly reduces offsets between amplifiers, but introduces read noise to each column.
{\it Fourth Panel:} Background subtraction with the mean value of the pixels below Y=15 and above Y=190 eliminates offsets between amplifiers without introducing any substantial new noise source.
}\label{fig:ampOffsetsOtisGrismSlope}
\end{figure*}

\section{1/f Noise}
The readout circuitry for NIRCam and other JWST NIR instruments introduces correlated noise from pixel-to-pixel.
This is visible when examining the time series of pixels as they are read out within a frame in the order they are addressed.\footnote{See a diagram of the readout directions at \url{https://jwst-docs.stsci.edu/near-infrared-camera/nircam-instrumentation/nircam-detector-overview/nircam-detector-readout}}
The power spectrum of this pixel-by-pixel time series reveals high power at low frequencies that drops approximately like a 1/f power law in the power spectrum of the time series.
Figure \ref{fig:pixelTSeriesPSpec} shows the average Lomb Scargle periodogram from amplifier over 108 dark frames.
To create the periodogram, the reference pixels are used to remove amplifier offsets between frames.
Next, the average of all frames is subtracted to remove the bias and all pixels that deviate by more than 80 DN from the average are masked to eliminate bad pixels.

@@

We note that Figure \ref{fig:genericSourceErrorContrib} shows a somewhat counter-intuitive result: read noise dominates for bright source and becomes less important for faint sources.
For typical observations, such as faint galaxies or distant stars, the opposite is true: read noise is more of a concern for faint sources than bright ones.
So why is there a discrepancy?
The difference in these two science cases is that for exoplanet transit observations, the detector well depth is filled to approximately the same level ($\sim 32,000$ DN or 50\% well depth per pixel) regardless if the system's magnitude is 6 or 11.
For the bright source with a few groups up the ramp, the read noise will be a larger fraction of 32,000 DN than for a faint source with many groups up the ramp.
This is different from a faint galaxy in the $K > 22$ range, which will not fill the detector well.
For a fixed exposure time, a $K=25$ and $K=20$ galaxies would be observed with approximately the same number of groups per integration so the read noise in DN from each source in an integration would be approximately the same whereas the number of DN from the source per integration would be 100 times larger for the $K=20$ galaxy.

@@

\begin{figure*}[!hbtp]
\centering
\includegraphics[width=.99\columnwidth]{fringing_grism_cv3.pdf}
\caption{A CV3 grism spectrum over the time series grism test shows evidence for fringing and where the oscillations exceed the photon and read noise (green point).
An illustrative sinusoidal function is plotted for comparison to better see how the frequencies change as a function of wavelength (X pixel).
}\label{fig:CV3GrismSpecFringing}
\end{figure*}

Figure \ref{fig:cv3ExtractedSpectrum} shows an extracted spectrum from Cryogenic Vacuum 3 testing at NASA Goddard where a broadband point source is dispersed by the grism.
The slope image is calculated from the raw ramp using the internal NIRCam ramps-to-slopes pipeline \texttt{ncdhas}.
The background is subtracted with a 2nd order polynomial fit and the flux is extracted with a variance-weighted summation \citep[e.g.][]{horne1986optimalE}.
The spectrum is normalized by a 3rd order spline fit with 400 knots to remove the broad shape of the spectrum and the throughput curves.
The resulting spectrum shows high frequency ($\sim$0.2 px$^{-1}$ to $\sim$0.5 px$^{-1}$) variations far in excess of the photon and read noise.
These oscillations increase in frequency at longer wavelengths (X pixels).
The variations have peak-to-valley amplitudes of about 8\%.
Figure \ref{fig:cv3ExtractedSpectrum} also shows an illustrative fringing model that increases in frequency as the wavelength decreases.

The fringing pattern is clearly visible in the grism spectrum, so it will be imprinted on all grism spectra.
However, this fringing pattern should be a function of the detector thickness and should remain constant for a fixed position in the detector.
The grism time series observations will be performed with target acquisition to reliably centroid on the source to better than 0.15 long wavelength pixels.
We therefore expect that the fringing pattern will be repeatable from one observation to the next.
For use in stellar characterization, however, it may be helpful to divide this fringing pattern out from a correction function made from a standard star.

\section{Subpixel Crosshatching}

@@

The crosshatch structure of the detectors extends down to the subpixel level, so it will not be fully corrected with a flat field division.
This sub-pixel structure has been imaged with microscopy on a Euclid HgCdTe detector \citep{shapiro2018crosshatch}.
Atomic force microscopy shows topological features that are approximately 1.2~$\mu$m in width \citep{chang2008surfaceMorphologyHgCdTe} compared to the 18~$\mu$m pixel sizes.
The width of the structures can also be estimated from the behavior for crosshatch lines as they cross pixel boundaries \citep{ninan2019crosshatchHPF}.
We estimate that the crosshatch pattern oriented at 90.9$\degree$ crosses 1 horizontal pixel for every 64 vertical pixels.
While crossing the boundary, there are $\approx 38$ rows where the crosshatch pattern spans more than 2 pixels.
If the crosshatch pattern has a tophat response function, then these 38 rows where the pattern spans two pixels imply a tophat full width of 0.6 pixels or a physical width of 10.8~$\mu$m for an 18~$\mu$m pixel pitch.
This is more than twice the estimate from \citet{ninan2019crosshatchHPF} for an HgCdTe used on the Habitable Planet Finder.
We expect that the crosshatch pattern's width varies among detectors or that a tophat function is a poor approximation of the actual subpixel response.
Within NIRCam detectors, there are large variations in the strength and orientation of crosshatch features.

The crosshatch pattern is wavelength dependent, as seen in Figure \ref{fig:crossHatchWavelengthDep}, where the throughput variations are the largest for short wavelengths (better resolving crystal structures) and smallest for the long wavelengths.
This is visible both in the throughput cross section of the flat field and the Fourier amplitude of the crosshatch as a function of wavelength.
We find a steeper wavelength dependence to the crosshatch pattern near 0.9$\degree$ from horizontal in the frequency domain or 0.9$\degree$ from vertical in the length domain.
%We also fit the crosshatch pattern.

@@

For the F300M filter and ALONG detector, we find 
for $\theta_1 = 0.906 \pm 0.001 \degree, \theta_2 = 68.605 \pm 0.001 \degree, \theta_3 = 113.075 \pm 0.001, a_1 = 2.774 \pm 0.002 \times 10^{-7}, a_2 = 1.569 \pm 0.002 \times 10^{-7}, a_3 = 1.744 \pm 0.002 \times 10^{-7}, b=0.345 \pm 0.0005 $ px$^{-1}$, $c=2.05 \pm 0.07 \times 10^{-2}$ px$^{-1}$, $d=6.78 \pm 0.01 10^{-3}$ and $e_b=0.26 \pm 0.01$ px$^{-1}$.
In other words, the separations between the vectors are $67.699 \degree, 67.831 \degree$, and $44.470 \degree$.
The uncertainties were simply derived from the diagonals of the covariance matrix and may be underestimated.

@@

NIRCam's 10 SCA's have very similar crosshatch angles with separations of 44.5 $\degree$ and 67.7 to 67.8 $\degree$ as seen in Figure \ref{fig:crosshatchModelGallery}.
They all have one primary crosshatch axis oriented either along the X pixels or the Y pixels, with two flanking axes at $\sim 67.7 \degree$ to either side.
The long wavelength arrays have a more pronounced crosshatch than the short wavelength arrays.
The NRCB4 array's primary crosshatch direction is most closely aligned with a pixel axis (the Y direction) than any SCA.
Figure \ref{fig:crosshatchModelGallery} shows the best-fit crosshatch 2D power spectra.

\subsection{Position of SubGrism Array}

@@

@@

The three subarrays considered are shown in Figure \ref{fig:subpixScanSimulation}.
We exclude reference pixels (which form a 4 pixel wide boundary around the detector) and an additional row on the bottom and top of the array because these illuminated rows are outliers in the flat field.
The first position is nearly the same as the SUBGRISM64 subarray position in the grism time series mode, except that the SUBGRISM64 includes the first 5 rows which are excluded here from the crosshatch analysis.
The field point used for the SUBGRISM64 mode is expected to be the same one as will be used for the SUBGRISM128 and SUBGRISM256 subarrays.
The position at the top of the ALONG detector has the smallest amplitude of crosshatch pattern in all three directions.
The best-fit amplitudes are used as input to the jitter simulations in Section \ref{sec:CrosshatchSim}.
We find that the maximum flux change for a 0.1 pixel shift is 118 ppm for the nominal grism position, 108 ppm for the top of the array and 153 ppm at the full frame grism position.
